---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading in the data


```{r}
library(tidyverse)
library(readr)
```


This function returns a list of boolean values indicating if the has digits (which it is supposed too) and using the summary function we can see the total number of cases. We are interested in the "FALSE" cases, and looking at the summary nearly.

For now my analysis will have to wait until the data can be fixed. With such a high number of cases manually cleaning each case is out of the question. So I am going to explore either finding a better way to load the data, or a way to shift the columns back and add the performance dates back in.

```{r}
library(tidyverse)
tester <- read.table("TrackDiscusData.txt", sep = "%", quote="\"")
View(tester)
MensDisc <- tidyr::separate(tester, V1,
                         c("Empty", "Rank", "Distance", "Name", "Country", "Birthdate", "Worthless", "Location", "Date"),
                         sep = "\\s{2,}", ) %>%
  select(-Empty)
view(MensDisc)
```

Let's get a rough count of the countries of origin

```{r}
basicCountry <- MensDisc %>% distinct(Name, .keep_all = TRUE) %>% count(Country, sort = TRUE)
ggplot(basicCountry, aes(x = Country, y = n, color = Country)) + 
  geom_col() +
  scale_x_discrete(guide = guide_axis(n.dodge = 5)) +
  labs(y= "Number of Athletes", x = "Country Code")
```
This is just a preliminary plot to give me an idea of how the distribution of the athletes country of origin. What the graph shows is that some countries clearly produce a majority of winning atheletes while other countries just have a few champions

Going forward I would like to investigate athletes countries year by year in separate graphs to see if certain countries started producing more or less champions over time

```{r}
ForOther <- MensDisc %>% mutate(Year = substr(Date,7,10))
```

Now I have a new row that is just the year with the rest of the date stripped away so we can do a quick visualization to see how athlete performance changed over time

```{r}
library(lubridate)

Bday   <-parse_date_time2(ForOther$Birthdate, "dmy", cutoff_2000 = 15)
Rday <-parse_date_time2(ForOther$Date, "dmY", cutoff_2000 = 20)
Dif <-difftime(Rday, Bday, units = "weeks")
Please <- as.integer(Dif/52)
New <- ForOther %>% 
  mutate(NewCol = as.integer(Please)) %>%
  mutate(Distance = as.numeric(ForOther$Time))
view(New)



```

Lets try to graph age and performance data!

```{r}
ggplot(New, aes(x=NewCol, y=Time, color=Country)) + geom_point(alpha=0.3) + scale_y_discrete(guide = guide_axis(n.dodge = 7))
```

 Looking at the scatterplot it is evident that the best 


